Benchmarking on Rigetti architecture

### 2022-08-10 txt/manuscript.tex
changes  pictures and proofs

In the scope of this Paper, we aim at  exploring ideas revolving around quantum 
gate model-inspired computing devices, and assess their feasibility to 
benchmark a modern quantum architectures. The main idea that we have envisioned 
for this project is  introducing new concepts of algorithms for NISQ devices 
benchmarking. This project aims to characterize the computing power and 
investigate possible practical applications of such devices having access by 
Amazon Braket.  First of all, we will focus on constraining quantum algorithms 
being relatively easy to implement on quantum architectures with 
straightforward, intuitive operational interpretation. Moreover, we need to 
possess a scalability, that is the ability to adjust to large range of 
architectures size.  
These properties will give us major advantage over the existing approaches  of 
benchmarking.

This idea is built upon two specific tasks that are to be achieved within this 
proposal. 
The first task is to   desing a benchmark based on discrimination of von 
Nemaumm measurements scheme. While the second one is to extension  this 
approach to benchmark on the grounds of certification of von Neumann 
measurements scheme.

\subsubsection{Model specification}
Theoretically, the scenario of this set-up assumes that Alice and Bob have an 
unknown measurement device, a black--box. Alice has hidden one of the two 
measurements, say $\SS$ and $\TT$,  in the black box. Bob only has information that it 
performs one of such measurements. 
First, they want to tell whether it is possible to discriminate $\SS$ and $\TT$ perfectly, i.e. with probability equal to one. If this is not the case, we would like to
know the upper bound of such a probability. 
Theoretical results of discrimination of 
von Neumann measurements are well-known and characterized 
in~\cite{puchala2018strategies}.   
They devised the optimal discrimination strategy between two von Neumann measurements and also calculated the exact probability of their discrimination.

In this work we will consider the problem of discrimination between the von Neumann measurement in the computational basis  and  the von Neumann measurement in basis $U$, where $U$ is a unitary matrix. 
This scheme is 
presented in Fig.~\ref{fig:scheme} and 
can be divided into the following steps:

\begin{figure}[h!]
	\centering 
	\includegraphics[width=0.9\linewidth]{gates1.pdf} 
	\caption{A schematic representation of the setup for distinguishing
		measurements by using entanglement. }
	\label{fig:scheme}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.5]{pics/theoretical_scheme}
	\caption{Theoretical discrimination scheme}
	\label{fig:theoretical_scheme}
\end{figure}
\todo[inline]{tan powinien byc $\PP_{V_i}$}

\begin{enumerate}\item Alice and Bob share entanglement state $\ket{\psi_{AB}}$.
	\item Alice performs one of two known measurements,  say $\PP_{\Id}$ and $\PP_{U}$,  on part $A$ of the input 
	state  $\ket{\psi_{AB}}$.
	\item Alice  uses
	the output label $i$ and performs the conditional binary measurement on the part 
	$B$.
	\item  Bob measures the part $B$ of the input state  $\ket{\psi_{AB}}$ and 
	makes a decision based on received label $j$. If $j=0$, he decides that the measurement $\PP_U$ is hidden in the black box. Otherwise, he decides that the measurement $\PP_{\Id}$ was hidden in the black box. 
\end{enumerate}

@@

Let us take a closer look at the setup for distinguishing von Neumann 
measurements (see Fig.~\ref{fig:scheme}). One of the components of such a 
scheme is to perform the conditional binary measurement. The first limitation 
is a lack of possibility of implementing the conditional binary measurement. 
The only possible measurement performed on quantum architectures is the measurement 
on a computational basis. 
Here, we present two possible methods that provide a clever way to cross the obstacle (two schemes of practical realization).

The first way is a postselection scheme which can be described as 
follow.  The schematic representation of this setup we can see on Fig.~\ref{postselection}.     
\begin{scheme}(By using postsellection)

\todo[inline]{PAMIETAJ O DAGGERACH  no i u nas to bedzie nazwane $\PP_{V^\dagger_j}$!!!} 
\begin{figure}[h!]
	\includegraphics[scale=1.5]{onequbit.pdf}
	\caption{A schematic representation of the setup for distinguishing
		measurements using postselection.}
	\label{postselection}
\end{figure}

\begin{figure}[h!]
	\centering 
	\includegraphics[scale=1.2]{pics/postselection}

	\caption{ A schematic representation of the setup for distinguishing
		measurements using postselection. 
	}\label{fig:postsellection}
\end{figure} 
\begin{enumerate}
	\item We prepare input state $\ket{\psi_{AB}}$.
	\item One of two measurements is performed on part A.
	\item We prepare optimal measurements $V_j$, which are well known by 
	Holevo--Helstrom theorem and perform them on part $B$.
	\item We receive an output label $i$.
	\item We include these cases for which $i = j$. Otherwise, we reject them.
	\item By the use of $V_j$ output, we decide whether the measurement was 
	performed on part $A$.
	\item We count the probability of postselection.
\end{enumerate}

%	\begin{enumerate}
%		\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
%		\rangle. $
%		\item We prepare one of two unitary channel $\Phi_{U} $ or $\Phi_{\1}$. 
%		\item We implement unitary $V_0 $ or $ V_1$.
%		\item We prepare the measurement $\Delta$ in standard basis (already exists on 
%		Rigetti architecture) on each qubits.
%		\item We calculate the probability of correct discrimination in the following 
%		way
%		\begin{equation}
%		p = ...
%		\end{equation}
%		%\todo[inline]{grubsza kmina, musze Ci to na zywo opowiedziec xd}
%	\end{enumerate}
\end{scheme}

\begin{scheme}(By using controlled unitary)
	\begin{figure}[h!]
		\centering 
		\includegraphics[scale=1.2]{pics/controlled_unitary}

		\caption{ A schematic representation of the setup for distinguishing
			measurements using controlled unitary gate. 
		}\label{fig:controlled}
	\end{figure}
	\begin{enumerate}
		\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
		\rangle. $
		\item We prepare one of two unitary channel $\Phi_{U} $ or $\Phi_{\1}$. 
		\item We implement controlled unitary $V_0 \oplus V_1$.
		\item We prepare the measurement $\Delta$ in standard basis (already exist on 
		Rigetti architecture) on each qubits.
		\item We calculate the probability of correct discrimination in the following 
		way
		\begin{equation}
		p = \frac{|j=0 \wedge \Phi_{U} | + |j=1 \wedge \Phi_{\1}|}{\text{all 
				cases}}
		\end{equation}
	\end{enumerate}
\end{scheme}

@@

Let $M_{d_1,d_2}$ be the set of all matrices of dimension $d_1 \times d_2$ over
the field $\mathbb{C}$. For  simplicity notation, square matrices will be 
denoted by
$M_d$.  The subset of $M_d$ consisting of Hermitian matrices of dimension $d$ 
will  be  denoted  by $\HH_d$,  while  the  set  of  positive semidefinite 
matrices of dimension $d$ by $\HH_d^+$. The set of quantum states, that is 
positive semidefinite operators having
trace equal to one, will be denoted $\Omega_d$.  The subset of $M_d$ consisting of 
unitary matrices will be denoted by
$\UU_d$, while its subgroup of diagonal unitary operators will be denoted by
$\DD \UU_d$.  We will also need a linear
mapping transforming $M_{d_1}$ into $M_{d_2}$, which will be denoted \begin{equation}
\Phi: M_{d_1 } \rightarrow M_{d_2}.
\end{equation}

Finally, we introduce a special subset of all mappings $\Phi$, called quantum channels, which are completely positive
and trace preserving (CPTP).
A unitary channel
$\Phi_{U}$ is defined as $\Phi_U(\cdot) = U \cdot U^\dagger$ for  $U \in 
\UU_d$. A general quantum
measurement, that is a positive operator valued measure (POVM) $\PP$ is a
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
\emph{effects}, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$. If
all the effects are rank-one projection operators, then such a measurement is
called a von Neumann measurement. Every von Neumann measurement can be
parameterized by a unitary matrix and hence we will use the notation $\PP_{U}$
for a von Neumann measurement with effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
where $\ket{u_i}$ is the $i+1$-th column of the unitary matrix $U$. The action of
quantum measurement $\PP_{U}$ on some state $\rho \in \Omega_d$ can be
expressed as the action of a quantum channel
$
\PP_{U} : \rho \rightarrow \sum_{i=0}^{d-1} \bra{u_i} \rho \ket{u_i} \proj{i}.
$ Every von Neumann measurement  $\PP_{U}$ can be also rewritten as $\Delta 
\circ \Phi_{(UE)^\dagger}$, where $\Delta$ denotes the completely dephasing 
channel.

\subsubsection{Discrimination of a quantum operations}
This section defines a norm on the space of linear mappings, known as the diamond norm,
and establishes some of its properties. 
Next, we will also explain the precise connection between this
norm and the task of quantum operations discrimination.

Let us now consider a  linear mapping transforming  square matrices into square 
matrices 
that is $\Phi: M_{d_1} \to M_{d_2}$. 
 We define the diamond norm of a map $\Phi$ as 
\begin{equation}
\|\Phi\|_\diamond = \max_{\|X\|_1=1} \| \left(\Phi \otimes \1\right) (X) \|_1.
\end{equation}

@@

% The Appendices part is started with the command \appendix;
% appendix sections are then done as normal sections
% \appendix

% \section{}
% \label{}

% References:
% If you have bibdatabase file and want bibtex to generate the
% bibitems, please use
%
%  \bibliographystyle{elsarticle-num} 
%  \bibliography{<your bibdatabase>}

% else use the following coding to input the bibitems directly in the
% TeX file.

>>>>

In the scope of this Paper, we aim at  exploring ideas revolving around quantum 
gate model-inspired computing devices, and assess their feasibility to 
benchmark a modern quantum architectures. The main idea that we have envisioned 
for this project is  introducing new concepts of benchmarks for NISQ devices 
benchmarking. This project aims to characterize the computing power and 
investigate possible practical applications of such devices having access by 
Amazon Braket.

\subsubsection{Model specification}
Theoretically, the scenario of this set-up assumes that Alice and Bob have an 
unknown measurement device, a black--box. Alice has hidden one of the two 
measurements, say $\SS$ and $\TT$,  in the black box. Bob only has information that it 
performs one of such measurements. 
They want to know the upper bound of the   probability of correct discrimination between $\SS$ and $\TT$ and construct the optimal discrimination strategy between them to achieve the highest possible probability of discrimination. 
Theoretical results of discrimination of 
von Neumann measurements are well-known and characterized 
in~\cite{puchala2018strategies}.

In this work we will consider the problem of discrimination between von Neumann measurement in the computational basis  and  the von Neumann measurement in basis $U$, where $U$ is a unitary matrix. 
This scheme is 
presented in Fig.~\ref{fig:theoretical_scheme} and 
can be divided into the following steps:

%\begin{figure}[h!]
%	\centering 
%	\includegraphics[width=0.9\linewidth]{gates1.pdf} 
%	\caption{A schematic representation of the setup for distinguishing
%		measurements by using entanglement. }
%	\label{fig:scheme}
%\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/theoretical_scheme}
	\caption{Theoretical discrimination scheme}
	\label{fig:theoretical_scheme}
\end{figure}

\begin{enumerate}\item Alice and Bob share entanglement state $\ket{\psi_{AB}}$.
	\item Alice performs one of two known measurements,  say $\PP_{\Id}$ and $\PP_{U}$,  on part $A$ of the input 
	state  $\ket{\psi_{AB}}$.
	\item Alice   performs the conditional binary measurement on the part 
	$B$ based on
	the output label $i$. 
	\item  Bob measures the part $B$ of the input state  $\ket{\psi_{AB}}$ and 
	makes a decision based on received label $j$. If $j=0$, he decides that the measurement $\PP_U$ is hidden in the black box. Otherwise, he decides that the measurement $\PP_{\Id}$ was hidden in the black box. 
\end{enumerate}

@@

Let us take a closer look at the setup for distinguishing von Neumann 
measurements (see Fig.~\ref{fig:theoretical_scheme}). One of the components of such a 
scheme is to perform the conditional binary measurement. The first limitation of NISQ devices 
is a lack of possibility of implementing the conditional binary measurement. 
The only possible measurement performed on quantum architectures is the measurement 
on a computational basis. 
Here, we present two possible methods that provide a clever way to cross the obstacle (two schemes of practical realization).

The first way is to use a postselection whereas the second method is to apply a controlled unitary. The methods are described above.

\begin{scheme}(By using postsellection)

\todo[inline]{wyzej mamy bez daggera, trzeba to ustalic} 
%\begin{figure}[h!]
%	\includegraphics[scale=1.5]{onequbit.pdf}
%	\caption{A schematic representation of the setup for distinguishing
%		measurements using postselection.}
%	\label{postselection}
%\end{figure}

\begin{enumerate}
	\item Alice and Bob prepare entanglement state $\ket{\psi_{AB}}$.
	\item Alice performs one of two quantum channels, $\Phi_{\Id}$ and $\Phi_{U}$,  on part $A$ of the input 
	state  $\ket{\psi_{AB}}$.
	\item Alice measures the part $A$ of the input state  $\ket{\psi_{AB}}$ and receives an output label $i$.
	\item  
	From Holevo--Helstrom theorem, 
	 Bob performs the conditional binary measurement $\PP_{V_j^\dagger}$ on the part 	$B$.
	\item They include these cases for which $i = j$. Otherwise, they reject them.
	\item
	Bob measures the part $B$ of the input state  $\ket{\psi_{AB}}$ and 
	makes a decision based on received label $j$. If $j=0$, he decides that the measurement $\PP_U$ is hidden in the black box. Otherwise, he decides that the measurement $\PP_{\Id}$ was hidden in the black box. 
\end{enumerate}

The schematic representation of this setup we can see on Fig.~\ref{fig:postsellection}.

\begin{figure}[h!]
	\centering 
	\includegraphics[scale=1.7]{pics/postselection}

	\caption{ A schematic representation of the setup for distinguishing
		measurements using postselection. 
	}\label{fig:postsellection}
\end{figure}

PROBABILITY 
		\begin{equation}
		p = ...
		\end{equation}
		%\todo[inline]{grubsza kmina, musze Ci to na zywo opowiedziec xd}

\end{scheme}

\begin{scheme}(By using controlled unitary)

	\begin{enumerate}
			\item Alice and Bob prepare entanglement state $\ket{\psi_{AB}}$.
%		\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
%		\rangle. $
		\item Alice performs one of two quantum channels, $\Phi_{\Id}$ and $\Phi_{U}$,  on part $A$ of the input 
		state  $\ket{\psi_{AB}}$.
		\item Alice and Bob implement controlled unitary $V_0 \oplus V_1$ obtained from Holevo--Helstrom theorem.	
		\item Alice and Bob prepare the measurement $\Delta$ in computational basis on theirs systems.
		\item Bob	makes a decision based on received label $j$. If $j=0$, he decides that the measurement $\PP_U$ is hidden in the black box. Otherwise, he decides that the measurement $\PP_{\Id}$ was hidden in the black box.

	\end{enumerate}

\begin{figure}[h!]
	\centering 
	\includegraphics[scale=1.5]{pics/controlled_unitary}

	\caption{ A schematic representation of the setup for distinguishing
		measurements using controlled unitary gate. 
	}\label{fig:controlled}
\end{figure}

PROBABILITY

\begin{equation}
p = \frac{|j=0 \wedge \Phi_{U} | + |j=1 \wedge \Phi_{\1}|}{\text{all 
		cases}}
\end{equation}

\end{scheme}

@@

Let $M_{d_1,d_2}$ be the set of all matrices of dimension $d_1 \times d_2$ over
the field $\mathbb{C}$. For  simplicity notation, square matrices will be 
denoted by
$M_d$.  The subset of $M_d$ consisting of Hermitian matrices of dimension $d$ 
will  be  denoted  by $\HH_d$,  while  the  set  of  positive semidefinite 
matrices of dimension $d$ by $\HH_d^+$. The set of quantum states, that is 
positive semidefinite operators having
trace equal to one, will be denoted $\Omega_d$.  The subset of $M_d$ consisting of 
unitary matrices will be denoted by
$\UU_d$, while its subgroup of diagonal unitary operators will be denoted by
$\DD \UU_d$.  We will also need a linear
mapping transforming $M_{d_1}$ into $M_{d_2}$, which will be denoted \begin{equation}
\Phi: M_{d_1 } \rightarrow M_{d_2}.
\end{equation} 
There
exists a bijection between introduced linear mappings $\Phi$ and set of matrices $M_{d_1d_2}$,  known as the Choi-
Jamio{\l}kowski isomorphism.

Finally, we introduce a special subset of all mappings $\Phi$, called quantum channels, which are completely positive
and trace preserving (CPTP).
A unitary channel
$\Phi_{U}$ is defined as $\Phi_U(\cdot) = U \cdot U^\dagger$ for  $U \in 
\UU_d$. A general quantum
measurement, that is a positive operator valued measure (POVM) $\PP$ is a
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
\emph{effects}, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$. If
all the effects are rank-one projection operators, then such a measurement is
called a von Neumann measurement. Every von Neumann measurement can be
parameterized by a unitary matrix and hence we will use the notation $\PP_{U}$
for a von Neumann measurement with effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
where $\ket{u_i}$ is the $i+1$-th column of the unitary matrix $U$. The action of
quantum measurement $\PP_{U}$ on some state $\rho \in \Omega_d$ can be
seen as  a measure-and-prepare quantum channel as follows 
$
\PP_{U} : \rho \rightarrow \sum_{i=0}^{d-1} \bra{u_i} \rho \ket{u_i} \proj{i}.
$ It is worth noting that a given  von Neumann measurement  $\PP_{U}$ can be also rewritten as $\Delta 
\circ \Phi_{(UE)^\dagger}$, where $\Delta$ denotes the completely dephasing 
channel.

\subsubsection{Diamond norm}
This section defines a special norm on the space of linear mappings, known as the diamond norm,
and establishes some of its properties. 
Next, we will also explain the precise connection between the diamond
norm and the task of quantum operations discrimination.

Let us  consider a  linear mapping transforming  square matrices into square 
matrices 
that is $\Phi: M_{d_1} \to M_{d_2}$. 
 We define the diamond norm of a map $\Phi$ as 
\begin{equation}
\|\Phi\|_\diamond = \max_{\|X\|_1=1} \| \left(\Phi \otimes \1\right) (X) \|_1.
\end{equation}

\begin{equation}
\| \Phi  \|_\diamond =  \max_{\rho \in \Omega_d} \| \left( \Id \otimes \sqrt{\rho} \right) \mathcal{J}_\Phi  \left( \Id \otimes \sqrt{\rho} \right)  \|_1  \end{equation}

\subsubsection{Discrimination of a quantum operations}

@@

 \section{}
 \label{}

 References:
 If you have bibdatabase file and want bibtex to generate the
 bibitems, please use

  \bibliographystyle{elsarticle-num} 
  \bibliography{<your bibdatabase>}

 else use the following coding to input the bibitems directly in the
 TeX file.

 \bibliographystyle{ieeetr}
 \bibliography{rigetti}
